---
title: "Covr"
author: Jim Hester
date: July 20, 2015
output:
   ioslides_presentation:
     widescreen: true
vignette: >
  %\VignetteIndexEntry{Covr Exercises}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Background #
## What is it? #

  - Test Coverage tool
  - Mid-December 2014
  - Coverage
    - R Code
    - Compiled C/C++/Fortran Code

## Motivation ##

  - Know what tests to add
  - Gauge of package robustness
  - Other languages coverage tools
    - [gcov](https://gcc.gnu.org/onlinedocs/gcc/Gcov.html)
    - [JCov](https://wiki.openjdk.java.net/display/CodeTools/jcov)
    - [coverage](https://pypi.python.org/pypi/coverage)
    - [simplecov](https://rubygems.org/gems/simplecov)
  - Services
    - [coveralls.io](https://coveralls.io/)
    - [codecov.io](https://codecov.io/)

## Existing Alternatives ##

  - [R-coverage](http://r2d2.quartzbio.com/posts/r-coverage-docker.html) by Karl Forner
    - Modify R source and add instrumentation
    - Requires patching and recompiling R source
  - [testCoverage](https://github.com/MangoTheCat/testCoverage) by Tom Taverner, Chris Campbell, Suchen Jin were the two
    - Alternate parser
    - Complicated implementation
    - No S4 support
    - Limited output formats
    - Not simple to call

## Demo ##

  - R terminal
  - Shiny Application
  - Coveralls.io
  - Codecov.io

## Configuration ##

  - Exclusions
    - whole files
    - line
    - ranges
    - In-source and outside

# Implementation #
## Abstract Syntax Tree ##

  - How it works vignette
  - Walk the Abstract Syntax Tree
    - If a call with srcref
      - Add a trace function
      - Perform the call

----

```{r eval = FALSE}
modify_lang <- function(x, f, ...) {
  recurse <- function(y) {
    ## if (!is.null(names(y))) names(y) <- f2(names(y))
    lapply(y, modify_lang, f = f, ...)
  }

  if (is.atomic(x) || is.name(x)) {
    ## Leaf
    f(x, ...)
  } else if (is.call(x)) {
    as.call(recurse(x))
  } else if (is.function(x)) {
    formals(x) <- modify_lang(formals(x), f, ...)
    body(x) <- modify_lang(body(x), f, ...)
    x
  } else if (is.pairlist(x)) {
    ## Formal argument lists (when creating functions)
    as.pairlist(recurse(x))
  } else if (is.expression(x)) {
    ## shouldn't occur inside tree, but might be useful top-level
    as.expression(recurse(x))
  } else if (is.list(x)) {
    ## shouldn't occur inside tree, but might be useful top-level
    recurse(x)
  } else {
    stop("Unknown language class: ", paste(class(x), collapse = "/"),
      call. = FALSE)
  }
}
```

## Modify Calls ##

  - insert function without changing output?
  - Braces evaluate expressions, return last

```{r}
identical({ 1 + 2; 3 + 4 }, `{`(1 + 2, 3 + 4))
```

```{r, eval = FALSE}
`{`(count(), as.call(recurse(x)))
```

## Source References ##

  - Where in source a call is from?
  - srcref
    - `option(keep.source = TRUE)`
    - srcref attribute attached to each call

## Replacing functions ##

  - Replace all references with modified
  - `testthat::with_mock()`
    - C function replaces function pointer
    - retains original definition

## S4 Methods ##

  - Normal functions defined directly in package namespace
  - S4 methods are defined in an environment based on their generic

```{r eval = FALSE}
replacements_S4 <- function(env) {
  generics <- getGenerics(env)

  unlist(recursive = FALSE,
    Map(generics@.Data, generics@package, USE.NAMES = FALSE,
      f = function(name, package) {
      what <- methodsPackageMetaName("T", paste(name, package, sep = ":"))

      table <- get(what, envir = env)

      lapply(ls(table, all.names = TRUE), replacement, env = table)
    })
  )
}
```

## Compiled Code ##

  - Gcov
    - built into gcc and clang
      - `-fprofile-arcs -ftest-coverage`
      - `-O0`
    - Need to override default and package makevars
      - `PKG_CFLAGS` puts optimization _before_ default `-O2`
    - Temporarily Modify `~/R/Makevars`
    - No results until process terminated
      - Call R subprocess

## Running Tests ##

  - `testthat::source_dir()`

```{r, eval = F}
lapply(dir("tests"), sys.source, envir=new.env(parent=emptyenv()))
```
## Sending Results ##

  - Track coverage over time
  - Report on coverage drops from contributions.
    - Codecov.io
    - Coveralls.io

  - Expect a JSON file
    - coverage per line
    - source code per line
  - Travis CI
    - Travis Job ID

## Future ##

  - Functions defined at run-time
  - if/else parsing
