---
title: "Covr Exercises"
author: "Jim Hester"
date: "`r Sys.Date()`"
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Covr Exercises}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

1. See how covr modifies the function calls to track coverage using `trace_calls()` on the function definition.
```{r}
fun1 <- function(x) {
  x + 1
}

covr:::trace_calls(fun1)
```

2. What happens when you call `trace_calls()` on a function with a conditional?
```{r}
fun2 <- function(x) {
  if (x > 5) {
    x + 2
  } else {
    x
  }
}

covr:::trace_calls(fun2)
```

3. What happens if that conditional does not use braces?  Why do you think this is?
```{r}
fun3 <- function(x) {
  if (x > 5)
    x + 2
  else
    x
}

covr:::trace_calls(fun3)
```

(Anyone interested in the in-progress fix for this see
[jimhester/covr#39](https://github.com/jimhester/covr/issues/39) and
[jimhester/covr@if_else](https://github.com/jimhester/covr/tree/if_else),
Suggestions welcome!).

4. What about functions that produce functions (closures)?
```{r}
fun4 <- function(x) {
  function(y) {
    x + y
  }
}

covr:::trace_calls(fun4)
```

5. Now lets try running `function_coverage()` to actually report on the
coverage.  Why is no filename displayed after the `Coverage:` line?

```{r}
cov <- function_coverage(fun1)
cov
```

Grouping by function rather than file makes more sense in this context.  It is
also useful to coerce the result into a `data.frame`.  Why is our coverage 0?
```{r}
print(cov, group = "function")

as.data.frame(cov)
```

You can also use `shine()` to start a shiny app to browse the results.
```{r}
shine(cov)
```

6. You can run any number of test commands as additional arguments.
```{r}
cov <- function_coverage(fun1, fun1(1))

as.data.frame(cov)
print(cov, group = "function")
```

7. Try doing a few more calls and verify the coverage value responds correctly.
```{r}
cov <- function_coverage(fun1, fun1(1), fun1(10))

as.data.frame(cov)
print(cov, group = "function")
```

8. What happens with a conditional? How about `fun3()`, which does not have braces.
```{r}
# true condition if x is less than 5
cov <- function_coverage(fun2, fun2(1))
as.data.frame(cov)

cov <- function_coverage(fun2, fun2(6))
as.data.frame(cov)

cov <- function_coverage(fun2, fun2(1), fun2(6))
as.data.frame(cov)
```

9. `package_coverage()` can be used to automatically run tests for a package.  Test using your package of choice (Examples use the new SummarizedExperiment package)
```{r, eval = F}
# In shell clone the code from the Bioconductor-mirrors
# git clone https://github.com/Bioconductor-mirror/SummarizedExperiment.git
setwd("~/SummarizedExperiment")

# then you need to install the package to make sure all of the dependencies are installed
BiocInstaller::biocLite("SummarizedExperiment")

# default runs the tests for the package
cov <- package_coverage()
```

Like before default print method prints per file
```{r, eval = FALSE}
cov
```

But you can also print per function
```{r, eval = FALSE}
print(cov, group = "function")
```

Or output a `data.frame`
```{r, eval = FALSE}
as.data.frame(cov)
```

`shine()` is more useful for browsing package coverage.
```{r, eval = FALSE}
shine(cov)
```

10. If your package does not have tests, covr can still be useful to show
vignette or example coverage. e.g. `type = "example"`
```{r, eval = FALSE}
cov <- package_coverage(type = "example")
```

11. You can run all of test, example and vignette coverage using `type =
"all"`.  In this case `shine()` provides a toggle button to switch between the
three types of coverage.
```{r, eval = FALSE}
cov <- package_coverage(type = "all")

shine(cov)
```
